Theoretical prediction of disrupted Min oscillation in flattened \emph{Escherichia coli}

### 2015-06-26 paper/response-to-referees.txt

Minor points:

4. The referee points out that MinD clusters had been considered by
Kruse prior to Huang.  We thank the referee for pointing out this
oversight, and have updated our background section to include this
reference and more accurately reflect the historical development.

5. Based on the referee's suggestion, we have clarified that reference
11 does not extend Huang's model, except in that it simulates the same
reaction-diffusion equations stochastically.

6. The referee points out that we failed to mention that stochastic
simulations are needed to explain the experimentally observed
length-dependence of fluctuations in the Min system.  We have added
mention of this additional evidence of the importance of inclusion of
stochastic effects, in the paragraph suggested by the referee.

### 2015-06-26 paper/plos.tex
answer one more minor objection Tests skipped!

{
  \color{red}
  Referee 1.
  \begin{enumerate}
  \item Need to better justify use of Huang.
  \item Couldn't another deterministic model predict Mannik's results?
  \item The stadia seem on average to be symmetric.
  \item MINOR:
  \item New citation for earlier observation.
  \item A couple of clarifications of our descriptions of cited papers.
  \item Stochastic are also important for a PNAS 2010 work.
  \end{enumerate}

  Referee 3
  \begin{enumerate}
  \item Skepticism about the effect of flattening, in terms of
    coherence times in particular.
  \item Want more clarity with regard to how we divide up the cell to
    compute coherence times.
  \item Starry night patterns are inconsistent with experimetns.
    Membrane diffusion may alleviate this problem.
  \item MINOR:
  \item Clarification in caption.  I don't quite understand.
  \item Figure 2 units for colormap.
  \item Figure 4 caption needs some title for the figure, not just
    ``We display here...''.
  \item ``coli'' should not be capitalized.
  \item FtsZ polymer does not form on the cell wall but near the
    cytoplasmic side of the inner membrane.
  \end{enumerate}
}

@@

There are only on the order of a thousand MinD proteins in a given
cell, which is fewer than the number of grid points used in our
simulations, suggesting that the deterministic continuum
description--which requires fractional numbers of proteins at each
grid point---may break down, and that stochastic behavior may play a
significant role.  To treat this, there have been a number of
applications of the Huang 2003 model that stochastically simulate the
same reaction-diffusion equations~\cite{fange2006noise,
  kerr2006division}.  These studies largely confirm the results of
Huang's deterministic model when applied to the wild-type, pill-shaped
phenotype.  However, the stochastic models are slightly more
successful in predicting experimentally observed oscillations in round
cell phenotypes~\cite{fange2006noise, huang2004min}, and they enable
prediction of fluctuations in the predicted
behavior~\cite{kruse2007experimentalist}.  Both deterministic and
stochastic models are widely used throughout systems
biology~\cite{lawson2013spatial, robb2014stochastic,
  oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
have unique advantages and limitations.  The deterministic approach
has an advantage in providing a simple prediction of average behavior,
while the stochastic approach enables prediction of fluctuations from
that mean, and reduces sensitivity to initial conditions.

>>>>

@@

There are only on the order of a thousand MinD proteins in a given
cell, which is fewer than the number of grid points used in our
simulations, suggesting that the deterministic continuum
description--which requires fractional numbers of proteins at each
grid point---may break down, and that stochastic behavior may play a
significant role.  To treat this, there have been a number of
applications of the Huang 2003 model that stochastically simulate the
same reaction-diffusion equations~\cite{fange2006noise,
  kerr2006division}.  These studies largely confirm the results of
Huang's deterministic model when applied to the wild-type, pill-shaped
phenotype.  However, the stochastic models are slightly more
successful in predicting experimentally observed oscillations in round
cell phenotypes~\cite{fange2006noise, huang2004min}, and they enable
prediction of fluctuations in the predicted
behavior~\cite{kruse2007experimentalist}, including the observation of
stochastic switching in short cells instead of a transition to static
behavior as would be predicted from a deterministic
model~\cite{fischer2010intra}.  Both deterministic and
stochastic models are widely used throughout systems
biology~\cite{lawson2013spatial, robb2014stochastic,
  oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
have unique advantages and limitations.  The deterministic approach
has an advantage in providing a simple prediction of average behavior,
while the stochastic approach enables prediction of fluctuations from
that mean, and reduces sensitivity to initial conditions.
